Combining information from photometric and spectroscopic data:\\ Probabilistic spectrophotometric inference

### 2015-07-21 tex/ms.tex
support for mocks with differnt run parameters, add CMD based values to joint real posterior plot

The calibration-vector and Gaussian-Process modified log-likelihood is
\begin{eqnarray}\label{eq:spectroscopicLF_v2}
\ln p(d\given\theta,\alpha) &=&
                                -\frac{1}{2}\,\left[\transpose{\Delta}\,C^{-1}\,\Delta
                                + \ln([2\pi]^N\,\det{C}) \right]
\\
\Delta &\equiv& d - \mu(\theta) \ast f(\alpha) 
\\
C_{nn'} &\equiv& [\sigma_n^2 + (\hat{s} \, \mu_n)^2]\,\delta_{nn'} +
 K_\alpha(\lambda_n - \lambda_{n'}) 
\quad ,
\end{eqnarray}
where $\ast$ denotes element-wise multiplication,
$\theta$ are the physical parameters,
$\alpha$ is a blob of nuisance parameters
(which determine the jitter, the calibration vector and the kernel function),
$\Delta$ is a residual column vector,
$f(\alpha)$ is a calibration column vector,
$C$ is the Gaussian Process covariance matrix,
made up of $N^2$ values $C_{nn'}$,
$K_\alpha$ is a kernel function that sets the wavelength scale and
amplitude of the residual calibration issues,
and each $\lambda_n$ is the wavelength of spectral pixel $n$.
The calibration vector $f(\alpha)$ is going to take care of the
large-scale calibration residual; the Gaussian Process is going to
pick up smaller-scale residuals away from \emph{that}.  The
flexibility of the Gaussian Process specified by $C$ will obviate the
need for numerous high order polynomial terms in $f(\alpha)$ and thus
reduce the number of free parameters required in the model.
This---equation~(\ref{eq:spectroscopicLF})---is the log-likelihood we
are going to use for the spectroscopic data in what follows.

@@

The two log-likelihoods, for photometry and spectroscopy, are combined
into a single likelihood function,
\begin{eqnarray}\displaystyle ln p(d_s, d_p\given \theta, \alpha, b) =
ln p_{spec}(d_s\given \theta, alpha) + ln p_{phot}(d_p\given \theta,
b) \quad .
\end{eqnarray} where now $d_s$ is the spectroscopic data and $d_p$ is
the photometric data.  The spectroscipc likelihood $p_{spec}$ is given
by equation~(\ref{eq:spectroscopicLF}) and $p_{phot}$ is given by
equation~(\ref{eq:photometricLF}).

@@

\item \sout{Mock perfectly calibrated data: inference using spectroscopy and one
  photometric band, assuming no calibration uncertainty.  This is as
  good as we could get given the data and its uncertainties.}

@@

\item Compare the physical parameters inferred from from the real
  calibrated and uncalibrated spectrum with each other and with CMD
  based parameters.

@@

We have demonstrated, using using both mock and real spectra of
galactic stellar clusters, that we can combine photometric and
spectroscopic data in a principled way to infer physical properties.
We do not resort to an arbitrary down-weighting or up-weighting of
certain data. Instead, we explicitly and flexibly model our uncertain
knowledge of the spectral calibration and infer this calibration
jointly with the physical properties. For a number of applications
this combination of spectroscopy and photometry reduces the need for
spectrophotometric calibration entirely, such that the less expensive
photometric calibration is sufficient.  Additional calibration
information, if it exists, can be incorporated into informative priors
on the calibration parameters of the model, inform the construction of
the calibration model (including the form of the covariance kernel),
or approximately remove gross features that are difficult to model
parametrically or via a Gaussian Process. A Guassian Process can be
used to model covaraint noise resulting from errors in an initial
calibration (of an amplitude $a$) thus propogating calibration
uncertainty into the model parameters $\theta$. 
{\color{blue} Discuss more complicated covariance kernels, multiple
kernels, combining fractional and absolute covariance amplitudes.}

{\color{blue} BDJ words (and experiments?) about the gains that might
be made by including photometry over a large range in wavelength with
spectroscopy over relatively small ranges.}  In this case even with
perfectly known spectrophotometric calibration, the photometry would
in principle add information.  This would occur if models with very
similar spectra in the observed range had differing SEDs outside that
range.

@@

>>>>

The calibration-vector and Gaussian-Process modified log-likelihood is
\begin{eqnarray}
\label{eq:spectroscopicLF}
\ln p(d\given\theta,\alpha) &=&
                                -\frac{1}{2}\,\left[\transpose{\Delta}\,C^{-1}\,\Delta
                                + \ln([2\pi]^N\,\det{C}) \right]
\\
\Delta &\equiv& d - \mu(\theta) \ast f(\alpha) 
\\
C_{nn'} &\equiv& [\sigma_n^2 + (\hat{s} \, \mu_n)^2]\,\delta_{nn'} +
 K_\alpha(\lambda_n - \lambda_{n'}) 
\quad ,
\end{eqnarray}
where $\ast$ denotes element-wise multiplication,
$\theta$ are the physical parameters,
$\alpha$ is a blob of nuisance parameters
(which determine the jitter, the calibration vector and the kernel function),
$\Delta$ is a residual column vector,
$f(\alpha)$ is a calibration column vector,
$C$ is the Gaussian Process covariance matrix,
made up of $N^2$ values $C_{nn'}$,
$K_\alpha$ is a kernel function that sets the wavelength scale and
amplitude of the residual calibration issues,
and each $\lambda_n$ is the wavelength of spectral pixel $n$.
The calibration vector $f(\alpha)$ is going to take care of the
large-scale calibration residual; the Gaussian Process is going to
pick up smaller-scale residuals away from \emph{that}.  The
flexibility of the Gaussian Process specified by $C$ will obviate the
need for numerous high order polynomial terms in $f(\alpha)$ and thus
reduce the number of free parameters required in the model.
This---equation~(\ref{eq:spectroscopicLF})---is the log-likelihood we
are going to use for the spectroscopic data in what follows.

@@

The two log-likelihoods, for photometry and spectroscopy, are combined
into a single likelihood function,
\begin{eqnarray}\displaystyle 
\ln p(d_s, d_p\given \theta, \alpha, b) =
\ln p_{spec}(d_s\given \theta, \alpha) + \ln p_{phot}(d_p\given\theta, b) 
\quad .
\end{eqnarray} where now $d_s$ is the spectroscopic data and $d_p$ is
the photometric data.  The spectroscopic likelihood $p_{spec}$ is given
by equation~(\ref{eq:spectroscopicLF}) and $p_{phot}$ is given by
equation~(\ref{eq:photometricLF}).

@@

\item \sout{Mock perfectly calibrated data: inference using spectroscopy and one
  photometric band, assuming no calibration uncertainty.  This is as
  good as we could get given the data and its uncertainties.} 
 Make sure this is run for the linear model.

@@

\item \sout{Compare the physical parameters inferred from from the real
  calibrated and uncalibrated spectrum with each other} and with CMD
  based parameters.

@@

We have demonstrated, using using both mock and real spectra of
galactic stellar clusters, that we can combine photometric and
spectroscopic data in a principled way to infer physical properties.
We do not resort to an arbitrary down-weighting or up-weighting of
certain data. Instead, we explicitly and flexibly model our uncertain
knowledge of the spectral calibration and infer this calibration
jointly with the physical properties. For a number of applications
this combination of spectroscopy and photometry reduces the need for
spectrophotometric calibration entirely, such that the less expensive
photometric calibration is sufficient.  Additional calibration
information, if it exists, can be incorporated into informative priors
on the calibration parameters of the model, inform the construction of
the calibration model (including the form of the covariance kernel),
or approximately remove gross features that are difficult to model
parametrically or via a Gaussian Process. A Guassian Process can be
used to model covaraint noise resulting from errors in an initial
calibration (of an amplitude $a$) thus propogating calibration
uncertainty into the model parameters $\theta$.

{\color{blue} Discuss more complicated covariance kernels, multiple
kernels, combining fractional and absolute covariance amplitudes.}
In the present study we have used a single, relatively simple kernel
function to construct the covariance matrix.  It is possible to
include multiple kernels by simply adding the resulting covariance
kernels.  For example, one could add an additional kernel with a small
amplitude and wavelength scale to account for model imperfections
\citep[e.g.,][]{czekala15} or include covariance structure expected
from the data reduction process \citep[e.g., ][]{bolton10} which
would absolute (i.e. not a fraction of the mean model.)

{\color{blue} BDJ words (and experiments?) about the gains that might
be made by including photometry over a large range in wavelength with
spectroscopy over relatively small ranges.}  In this case even with
perfectly known spectrophotometric calibration, the photometry would
in principle add information.  This would occur if models with very
similar spectra in the observed range had widely differing SEDs
outside that range.

@@

\bibitem[Bolton \& Schlegel(2010)]{bolton10} Bolton, A.~S., \& 
Schlegel, D.~J.\ 2010, \pasp, 122, 248
